---
title: "Homework 6"
author: "Alex Hentz"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Google Colab Link: https://colab.research.google.com/drive/19O7TueX5x_Y0F0QPslXqu-ghP8eS5owS?usp=sharing

```{r}
library(tidyverse)
library(mdsr)
library(ggplot2)
library(plotly)
library(lubridate)
library(dplyr)
library(macleish)
what <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI401/main/data/whately_2015.csv")

what<-whately_2015
what$when <- as.Date(what$when, "%Y-%m-%d")

g<-what %>% 
  group_by(when) %>% 
  summarise(min_temp=min(temperature),
            max_temp=max(temperature),
            avg_temp=mean(temperature))
  
graph<-ggplot(g, aes(x = when)) +
  geom_line(aes(y = max_temp, color = "High Temp")) +
  geom_line(aes(y = avg_temp, color = "Avg Temp")) +
  geom_line(aes(y = min_temp, color = "Low Temp")) +
  labs(title = "Daily Temp Trends",
       x = "Time (in days)",
       y = "Temp") +
  scale_color_manual(values = c("High Temp" = "#FF6699", "Avg Temp" = "#99ccff", "Low Temp" = "#9999ff")) + theme_minimal()
ggplotly(graph)


