---
title: "DSCI 401 Project"
author: "Alex Hentz"
date: "2023-11-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(readxl)
library(plotly)
library(stringr)
federal_net_cost <- read_excel("C:/Users/alexh/OneDrive/federal_net_cost.xls")
View(federal_net_cost)

# Create a Bar Chart of all the agencies and their gross cost of 2022 and 2001
federal_net_cost$`Gross Cost (in Billions)` <- as.numeric(gsub("[^0-9.]", "", as.character(federal_net_cost$`Gross Cost (in Billions)`)))
federal_net_cost$`Subtotal (in Billions)` <- as.numeric(gsub("[^0-9.]", "", as.character(federal_net_cost$`Subtotal (in Billions)`)))

hidden_row <- federal_net_cost$`Agency Name` == "Total"
new_hidden_row <- federal_net_cost[!hidden_row, ]

gross_cost <- new_hidden_row %>% group_by(`Statement Fiscal Year`) %>% select(`Agency Name`, `Gross Cost (in Billions)`) %>% filter(`Statement Fiscal Year` == 2022)

bgc <- ggplot(data = gross_cost, aes(x = `Agency Name`, y = `Gross Cost (in Billions)`, fill = `Agency Name`)) +
  geom_bar(stat = "identity", width=1) + scale_x_discrete(label = function(x) stringr::str_trunc(x, 12)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("2022 Federal Government Gross Cost")
ggplotly(bgc)

# 2001 Graph

gross_cost_2 <- new_hidden_row %>% group_by(`Statement Fiscal Year`) %>% select(`Agency Name`, `Gross Cost (in Billions)`) %>% filter(`Statement Fiscal Year` == 2001)

bgcc <- ggplot(data = gross_cost_2, aes(x = `Agency Name`, y = `Gross Cost (in Billions)`, fill = `Agency Name`)) +
  geom_bar(stat = "identity", width=1) + scale_x_discrete(label = function(x) stringr::str_trunc(x, 12)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("2001 Federal Government Gross Cost")

ggplotly(bgcc)

# Create a chart of 2022 and 2011 of the subtotal cost of the agencies


subtotal <- new_hidden_row %>% group_by(`Statement Fiscal Year`) %>% select(`Agency Name`, `Subtotal (in Billions)`) %>% filter(`Statement Fiscal Year` == 2022)

bnc <- ggplot(data = subtotal, aes(x = `Agency Name`, y = `Subtotal (in Billions)`, fill = `Agency Name`)) +
  geom_bar(stat = "identity", width=1) + scale_x_discrete(label = function(x) stringr::str_trunc(x, 12)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("2022 Federal Government Subtotal Cost")

ggplotly(bnc)

subtotal_2 <- new_hidden_row %>% group_by(`Statement Fiscal Year`) %>% select(`Agency Name`, `Subtotal (in Billions)`) %>% filter(`Statement Fiscal Year` == 2011)

bncc <- ggplot(data = subtotal_2, aes(x = `Agency Name`, y = `Subtotal (in Billions)`, fill = `Agency Name`)) +
  geom_bar(stat = "identity", width=1) + scale_x_discrete(label = function(x) stringr::str_trunc(x, 12)) + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + ggtitle("2011 Federal Government Subtotal Cost")

ggplotly(bncc)


# President term spending comparison

Transportation <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Transportation") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Transportation)

Social_Security <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Social Security Administration") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Social_Security)

Defense <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Defense") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))  
print(Defense)

Education <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Education") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Education)

Energy <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Energy") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Energy)

Health_Human <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Health and Human Services") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Health_Human)

Veterans <- federal_net_cost %>% 
  mutate(Four_Year_Period = cut(`Statement Fiscal Year`, breaks = c(2000, 2004, 2008, 2012, 2016, 2020, 2024), labels = c("2001-2004", "2005-2008", "2009-2012", "2013-2016", "2017-2020", "2021-2024"), include.lowest = TRUE)) %>%
  filter(`Agency Name` == "Department of Veterans Affairs") %>%
  group_by(Four_Year_Period) %>% 
  summarise(Sum_Gross_Cost = sum(`Gross Cost (in Billions)`, na.rm = TRUE))
print(Veterans)


# Stacked Bar chart
combined_data <- rbind(
  mutate(Transportation, Agency = "Transportation"),
  mutate(Social_Security, Agency = "Social Security"),
  mutate(Defense, Agency = "Defense"),
  mutate(Education, Agency = "Education"),
  mutate(Energy, Agency = "Energy"),
  mutate(Health_Human, Agency = "Health and Human Services"),
  mutate(Veterans, Agency = "Veterans")
)

# Create a stacked bar chart
stack_dep <- ggplot(combined_data, aes(x = Four_Year_Period, y = Sum_Gross_Cost, fill = Agency)) +
  geom_bar(stat = "identity") +
  labs(title = "Federal Net Cost by Agency Over Time",
       x = "Four Year Period",
       y = "Sum Gross Cost (in Billions)",
       fill = "Agency") +
  theme_minimal()
ggplotly(stack_dep)